---
title: "Data358 Lab2: Notes"
subtitle: "Data splitting and Resampling methods"
author: "Taylor Stacy"
date: "`r Sys.Date()`"
output:
  html_document:
    number_section: no
    toc: no
    toc_depth: 3
    toc_float: yes
    code_folding: show
    css: lab_templet.css
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```
## Required packages
```{r}
library(AmesHousing)
library(rsample)
library(caret) #wrapper/megapackage
```
## Import Data
```{r}
ames <- AmesHousing::make_ames()
# :: are used to reference a function or package
glimpse(ames)
```
## Simple random Sampling
### Using 'caret' package
```{r}
set.seed(1234) #for reproducibility
# Step 1: create an index vector
inTrain <- createDataPartition(
  y = ames$Sale_Price,
  p = 0.7,
  list = FALSE)
  # specify outcome variable, % of data in training set, avoids returning the dataset)
# Step 2: Partition the data
train_1 <- ames[inTrain,]
test_1 <- ames[-inTrain,] # can also use setdiff() to create test set
print(train_1)
print(test_1)
```
### Using 'dplyr::sample_frac()' function
```{r}
set.seed(1234)
train_2 <- ames %>% 
  sample_frac(0.7)
test_2 <- ames %>% 
  setdiff(train_2)
print(train_2)
print(test_2)
```
### Using 'rsample::initial_split()' function
```{r}
set.seed(1234)
inTrain  <- initial_split(
  ames, prop = 0.7)
train_3  <- training(inTrain)
test_3   <- testing(inTrain)
print(train_3)
print(test_3)
```
## Import Credit Card Fraud Data
```{r}
ccfraud <- read_csv(here("DATA358","Data","ccfraud.csv"))

ccfraud %>% 
  mutate(Class = as.factor(Class)) %>% 
  group_by(Class) %>% 
  summarise(freq = n()) %>% 
  mutate(prop = freq/sum(freq))
```
## Stratified Random Sampling
### Using 'caret'package
```{r}
set.seed(1234)
inTrain <- createDataPartition(
  y = ccfraud$Class, p = 0.7, 
  list = FALSE)
stratified_train <- ccfraud[inTrain, ] 
table(stratified_train$Class) 
```
### Using dplyr::sample_frac()
```{r}
set.seed(1234)          
stratified_train <- ccfraud %>% 
  group_by(Class) %>%
  mutate(n_rows=n()) %>% 
  sample_frac(0.7, weight=n_rows) %>% 
  select(-n_rows) %>% 
  ungroup
table(stratified_train$Class) 
```
### Using rsample::initial_split()
```{r}
set.seed(1234)
inTrain  <- initial_split(
  ccfraud, prop = 0.7, 
  strata = "Class")
stratified_train  <- training(inTrain)
table(stratified_train$Class)
```
## Up Sampling
```{r}
set.seed(1234)
up_train <- upSample(
  x = ccfraud[, -ncol(ccfraud)],
  y = factor(ccfraud$Class))
table(up_train$Class) 
```
## Down Sampling
```{r}
set.seed(1234)
down_train <- downSample(
  x = ccfraud[, -ncol(ccfraud)],
  y = factor(ccfraud$Class))
table(down_train$Class)
```
## Leave-One-Out Cross Validation
```{r}
# Define training control
train.control <- trainControl(method = "LOOCV")
# Train the model
set.seed(1234)
model <- train(Sale_Price~Lot_Area,       # specify the model
               data = ames,               # specify data
               method = "lm",             # lm is for linear model
               trControl = train.control)
# Summarize the results
print(model)
```
## k-Fold Cross Validation
```{r}
# Define training control
train.control <- trainControl(method = "cv", 
                              number = 10) # specify k
# Train the model
set.seed(1234) 
model <- train(Sale_Price~Lot_Area,        # specify the model
               data = ames,                # sepecify data
               method = "lm",              # lm is for linear model
               trControl = train.control)
# Summarize the results
print(model)
```