---
title: "DATA358: Lab 09 Notes"
author: "Taylor Stacy"
date: "`r Sys.Date()`"
output:
  html_document:
    number_section: no
    toc: no
    toc_depth: 3
    toc_float: yes
    code_folding: show
    css: lab_templet.css
---

```{r setup, include=FALSE}
rm(list=ls())
library(caret)
library(cluster)
library(tidyverse)
library(corrplot)
library(gridExtra)
library(factoextra)
library(kableExtra)
library(here)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# Data
```{r}
customer <- read_csv(here("Data/Mall_Customers.csv"))
glimpse(customer)
```
### Unusual Values
```{r}
summary(customer[,3:5])
```
### Standardization
```{r}
features <- customer[,3:5]
features_z <- scale(features)
```

# k-means Clustering
## Elbow Method
```{r}
set.seed(1234)
fviz_nbclust(features_z,
             kmeans,
             method = "wss")
```

## Gap Statistic
```{r}
set.seed(1234)
fviz_nbclust(features_z,
             kmeans,
             method = "gap_stat")
```

## K-mean Clusters
```{r}
set.seed(1234)
customers_cluster <- kmeans(features_z,
                            centers= 6,
                            nstar=25)
names(customers_cluster)
```
### Visualizing the Cluster
```{r}
fviz_cluster(customers_cluster,
             data = features_z,
             ggtheme = theme_minimal(),
             show.clust.cent = TRUE,
             main = "Cluster of Clients")
```

### Characterizing the Clusters
```{r}
customer2 <- customer %>% 
  mutate(cluster = customers_cluster$cluster)
cust_summary <- customer2 %>% 
  group_by(cluster) %>% 
  summarise(n = n(),
            Mean_age = mean(Age),
            Mean_income = mean(Annual.Income),
            Mean_spending = mean(Spending.Score))
```
```{r}
gender <- customer2 %>% 
  group_by(cluster, Gender) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  select(-n) %>% 
  spread(Gender, prop)
gender
```

# Hierachical Clustering
```{r}
seeds <- read_csv(here('Data/seeds.csv'))
glimpse(seeds)
```
### Data Preparation
```{r}
# Define column names
cNames <- c("area", "perimeter",
            "compactness","length_kernel",
            "width", "asymmetry_coeff",
            "length_groove", "type")
#Apply column names.
colnames(seeds) <- cNames
# Create a data set of features
df <- seeds %>% 
  select(-c("compactness","type"))
```

### Data Exploration
```{r}
# Standardize the features
df = scale(df)
#View the first 6 rows for confirmation
round(head(df, nrow=6),4)
```

## Model Training
Two Approaches
  - Agglomerative clustering (used for this lab)
  - Divisive hierarchical clustering:

Four-steps
  1. Compute distance matrix:
  2. Use linkage function to group objects into hierarchical cluster tree
  3. Determining where to cut the hierarchical tree into clusters
  4. Cut the hierarchical tree into clusters. This creates a partition of the data

### Distance Matrix
```{r}
# compute distances
res.dist <- dist(df, 
                 method = "euclidian")
# Visualize distances
fviz_dist(res.dist,
          lab_size = 5)
```

### Dendrogram
```{r}
res.hc.ward <- hclust(
  d = res.dist,
  method = "ward.D2")
# Visualize the dendrogram
fviz_dend(
  res.hc.ward,
  cex = 0.5,
  main="Cluster Dendrogram Using Ward's Method")+
  geom_hline(yintercept=20, 
             color="red", 
             linetype="dashed")
```

We used Ward's minimum variance linkage

### Determining Where to Cut the Tree
```{r}
# Cut in 3 groups and color by groups
fviz_dend(
  res.hc.ward, 
  k = 3,  # Cut in three groups
  cex = 0.5, # label size
  k_colors = c("#2E9FDF", "#4245f5", "#991a5a"),
  color_labels_by_k = TRUE, # color labels by groups
  rect = TRUE, # add rectangles around groups
  ggtheme = theme_classic()     # Change theme
  )
```

### Visualize Tree
```{r}
# Cut tree into 3 groups
grp <- cutree(res.hc.ward, k = 3)
# Add clusters to data set
seeds_new<-as.data.frame(seeds)
seeds_new$cluster<-grp

fviz_cluster(
  list(data = df, cluster = grp),
  palette = c("#2E9FDF", "#4245f5", "#991a5a"),
  ellipse.type = "convex", # Concentration ellipse
  show.clust.cent = FALSE, 
  ggtheme = theme_minimal()
)
```