---
title: "data 358 lab 6 notes"
subtitle: "Classification Methods: LDA and QDA"
author: "Taylor Stacy"
date: "`r Sys.Date()`"
output:
  html_document:
    number_section: no
    toc: no
    toc_depth: 3
    toc_float: yes
    code_folding: show
    css: lab_templet.css
---

```{r setup, include=FALSE}
library(tidyverse)
library(caret)
library(rsample)
library(MASS)
library(here)
library(dplyr)
library(reshape2)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
# clears environment
rm(list = ls())
```

# Data
```{r}
vehicle <- read_csv(here("Data/vehicle3.csv"))
str(vehicle)
sapply(vehicle,class)
```
## Cleaning Data
```{r}
# remove ID variable and convert class to factor
vehicle <- vehicle %>% 
  dplyr::select(-1) %>% 
  mutate(Class = factor(Class))
table(vehicle$Class)
```

- 18 predictors: $p=18$ 
- 3 categories: $k=3$

# Data Exploration
## Missing Values
```{r}
anyNA(vehicle)
```
- No missing values

## Predictor Distributions
```{r}
# Create a box plot for all predictors
# reshape to long format

melt_df <- melt(vehicle[-19])

ggplot(melt_df, aes(x=variable, y=value))+
  geom_boxplot()+
  facet_wrap(~variable, scale = "free")
```

- There are few features with highly skewed distribution. Ideally we would transform these variables to make more symmetric

## Check equal variance
```{r}
summary(vehicle[-19])
```

##  Model Training
```{r}
set.seed(1234)
inTrain <- vehicle$Class %>% 
  createDataPartition(p =0.7,
                      list = FALSE)
train <- vehicle[inTrain,]
test <- vehicle[-inTrain,]
```

# Linear Discriminant Analysis (LDA) Model
- Use the `lsa()` function of the `MASS` package

## LDA: Model Training
```{r}
lda_fit <- lda(Class ~ ., data = train)
lda_fit
```
## Stat Histogram
```{r}
lda.values <- predict(lda_fit)
ldahist(data = lda.values$x[,1], g = train$Class)
```

## Scatterplot of discrimant functions
```{r}
# extract loadings
loadings <- data.frame((lda.values$x))
# add Class variable as factor
loadings  <- loadings %>% 
  mutate(Type = as.factor(train$Class))

# Plot LD1 and LD2
ggplot(data = loadings, aes(x = LD1,
                            y = LD2,
                            color = Type,
                            shape = Type))+
  geom_point()
```

## Prediction Matrix
```{r}
# Make predictions on test set
pred_test <- predict(lda_fit,test)

# extract prediction class
pred_class <- as.factor(pred_test$class)

confusionMatrix(pred_class, as.factor(test$Class))
```
# Quadratic Discriminant Analysis (QDA) Model
```{r}
qda_fit <- qda(Class ~ ., data = train)
qda_fit
```

## Confusion Matrix
```{r}
pred_test2 <- predict(qda_fit,test)
pred_class2 <- as.factor(pred_test2$class)

confusionMatrix(pred_class2, as.factor(test$Class))
```